Drought-sensitive trees survive on deep-water access in a tropical forest community

### 2020-08-30 manuscript.rmd
1. added fig for ERD species : obs+modelled plc curves 2. references to comm plc curves 3. removed suffix to WSG and LMA 4. removed references to trade-off figs.

## Model for Leaf Vulnerability Curves at Community-scale
\
We measured ($K_{leaf}$ vs. $\Psi_{leaf}$) on adult trees of `r length(sp.exp.param$sp)` common species (see [SI](#SIData)--RCT:"Joseph, please add a section"). We fitted four different functions to species-spefic data using the *anneal* optimization function with the package likelihood in R: exponential without an intercept ($K_{leaf} = Ae^{-B\Psi_{leaf}}$), exponential with an intercept ($K_{leaf} = C + Ae^{-B\Psi_{leaf}}$), logistic ($K_{leaf} = \frac{A}{1 + \left(\frac{\Psi_{leaf}}{x_{0}}\right)^B}$) and sigmoidal ($K_{leaf} = \frac{A}{1 + e^{-\left(\frac{\Psi_{leaf} - x_{0}}{B}\right)}}$) [@Guyot:2011]. The differences among AICc values for these functions were not large within species (Fig. S\@ref(fig:leaf-vcurves-data1) & S\@ref(fig:leaf-vcurves-data2)). We chose the exponential function without the intercept to fit observed data across species, and as a general function to scale-up to the community, because it had the lowest AICc for most species, captured the form of the observed data across all species well, and had the lowest number of parameters, that is, two:

\begin{equation}
K_{leaf} = Ae^{-B\Psi_{leaf}} + error
(\#eq:vc-exp)
\end{equation}
where $A$ and $B$ are fitted species-specific parameters. From these fits we estimated species $K_{max,~leaf}$, $\Psi_{50,~leaf}$ and  $\Psi_{80,~leaf}$ (see Table 1). We identified scaling relationships between fitted parameters $A$ and $B$ in Eq. \@ref(eq:vc-exp) and two traits that were available at the community-scale, namely,  WSG~100^$\circ$^C~ (hereafter, $WSG$, g cm^-3^) and LMA~lamina~ (hereafter, $LMA$; g m^-2^) (see SI for description of how data gaps were filled. RCT:"Joe, please add a reference if these data are published, else add an SI section."). We fitted polynomial equations described as,

@@

## Leaf Vulnerability Curves
\
For the leaf vulnerability curve parameters $A$ & $B$ fitted to the observed data on $K_{leaf}$ vs. $\Psi_{leaf}$, $WSG$ & parameter $B$ explained a large proportion of variation in parameter $A$ (Eq. \@ref(eq:a-mod); *Adj. R*^2^ = `r afit.r2`; *P* `r  afit.p`; also see Fig. S\@ref(fig:ab-fit)). $WSG$ & $LMA$ explained a large proportion of variation in parameter $B$ (Eq. \@ref(eq:b-mod); Adj. *R*^2^ = `r bfit.r2`; *P* `r  bfit.p`). WSG alone explained `r round(a.wsg.r2*100,0)` % of variation in $A$ ($A = `r a.wsg.cf[1]` `r a.wsg.cf[2]`B + `r a.wsg.cf[3]`B^2 + error$; *P* `r  a.wsg.p`) and `r round(b.wsg.r2*100,0)` % of variation in $B$ ($B = `r b.wsg.cf[1]` `r b.wsg.cf[2]`B + `r b.wsg.cf[3]`B^2 + error$; *P* `r  b.wsg.p`). Parameter $B$ alone explained `r round(a.b.r2*100,0)` % of variation in $A$ (*P* `r  a.b.p`), while $LMA$ alone was weakly related to $A$ and was not related to $B$ (not shown). The leaf vulnerability curves fitted to observed data on `r length(sp.exp.param$sp)` species are shown in Figs. S\@ref(fig:leaf-vcurves-data1), S\@ref(fig:leaf-vcurves-data2) and S\@ref(fig:obs-plc-curves) and those predicted for the `r length(unique(data.model.AB$sp))` species in the community are shown in Fig. \@ref(fig:comm-vuln-curves). Corresponding parameters for the community are given in Table S\@ref(tab:param) (also see Fig. \@ref(fig:obs-plc-curves)).

@@

```{r erd-hyd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-hyd}Modelled Effective Rooting Depth (ERD; mean $\\pm$ 1SE in m) versus hydraulic properties for seven large, canopy species found on Barro Colorado Island (Panama). *R*^2^ and *P*-values are given in panel insets.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/erd.stem.traits.jpeg"))
```

@@

### Future Directions for Model Development
\
The trade-offs that we identify between above- and below-ground traits, the water environment vertically below-ground and mortality rates are important for representing diversity in Dynamic Global Vegetation Models especially in species-rich forests. The hydraulic efficiency-safety trade-off between high values of maximum leaf or stem conductivity (high $K_{max}$) and cavitation-resistance (low $\Psi_{50}$ or $\Psi_{88}$) is hypothesized although community-wide examples are rare [@Hacke:2006; @Gleason.2016]. We report such a trade-off at both leaf (*n* = `r length(sp.exp.param$sp)` species) and stem level (*n* = `r length(hyd$KmaxS[!is.na(hyd$KmaxS)])` species; see Fig.S\@ref(fig:se-trade-offs) [RCT:"Nobby et al and Brett, are such trade-offs part of your other manuscripts? If not, we could consider moving these Figs. to the results section."]). We also found that LMA~lamina~ and WSG~100^$\circ$^C~ were strong predictors of PLC curves for leaves, but not for stems (not shown). Acquisitive traits such as high $K_{max}$ are expected to be associated with faster growth rates. We found a positive but non- significant relationship of adult growth rates ($\ge$ 10cm dbh) with $K_{max, ~leaf}$ (*R*^2^ = 0.13; *p* = 0.126, *n* = `r length(sp.exp.param$sp)`) and no relationship with $K_{max, ~stem}$ (*n* = `r stem.traits.sp` species). The relationships between ERD and above-ground hydraulic traits that we find will provide an important insight on how to model rooting depths and their coordination or trade-offs with other traits in order to better represent the functional diversity of tropical forests.

@@

>>>>

## Leaf Vulnerability Curves
\
We measured ($K_{leaf}$ vs. $\Psi_{leaf}$) on adult trees of `r length(sp.exp.param$sp)` common species (see [SI](#SIData)--RCT:"Joseph, please add a section"). We fitted four different functions to species-spefic data using the *anneal* optimization function with the package likelihood in R: exponential without an intercept ($K_{leaf} = Ae^{-B\Psi_{leaf}}$), exponential with an intercept ($K_{leaf} = C + Ae^{-B\Psi_{leaf}}$), logistic ($K_{leaf} = \frac{A}{1 + \left(\frac{\Psi_{leaf}}{x_{0}}\right)^B}$) and sigmoidal ($K_{leaf} = \frac{A}{1 + e^{-\left(\frac{\Psi_{leaf} - x_{0}}{B}\right)}}$) [@Guyot:2011]. The differences among AICc values for these functions were not large within species. We chose the exponential function without the intercept to fit observed data across species, and as a general function to scale-up to the community, because it had the lowest AICc for most species, captured the form of the observed data across all species well, and had the lowest number of parameters, that is, two:

\begin{equation}
K_{leaf} = Ae^{-B\Psi_{leaf}} + error
(\#eq:vc-exp)
\end{equation}
where $A$ and $B$ are fitted species-specific parameters. From these fits we estimated species $K_{max,~leaf}$, $\Psi_{50,~leaf}$ and  $\Psi_{80,~leaf}$ (see Table 1). We identified scaling relationships between fitted parameters $A$ and $B$ in Eq. \@ref(eq:vc-exp) and two traits that were available at the community-scale, namely,  WSG and LMA (@Wright:2010; see SI for description of how data gaps were filled. We fitted polynomial equations described as,

@@

## Leaf Vulnerability Curves
\
For the leaf vulnerability curve parameters $A$ & $B$ fitted to the observed data on $K_{leaf}$ vs. $\Psi_{leaf}$, $WSG$ & parameter $B$ explained a large proportion of variation in parameter $A$ (Eq. \@ref(eq:a-mod); *Adj. R*^2^ = `r afit.r2`; *P* `r  afit.p`; also see Fig. S\@ref(fig:ab-fit)). $WSG$ & $LMA$ explained a large proportion of variation in parameter $B$ (Eq. \@ref(eq:b-mod); Adj. *R*^2^ = `r bfit.r2`; *P* `r  bfit.p`). WSG alone explained `r round(a.wsg.r2*100,0)` % of variation in $A$ ($A = `r a.wsg.cf[1]` `r a.wsg.cf[2]`B + `r a.wsg.cf[3]`B^2 + error$; *P* `r  a.wsg.p`) and `r round(b.wsg.r2*100,0)` % of variation in $B$ ($B = `r b.wsg.cf[1]` `r b.wsg.cf[2]`B + `r b.wsg.cf[3]`B^2 + error$; *P* `r  b.wsg.p`). Parameter $B$ alone explained `r round(a.b.r2*100,0)` % of variation in $A$ (*P* `r  a.b.p`), while $LMA$ alone was weakly related to $A$ and was not related to $B$ (not shown). Leaf vulnerability curves fitted to observed data on `r length(sp.exp.param$sp)` species as well as percent loss of conuctivity curves are shown in Fig. S\@ref(fig:obs-plc-curves). Vulnerability curves were predicted using Eqs. \@ref(eq:a-mod)-\@ref(eq:b-mod) for `r sp.n.model` species for which such curves were unavailable (out of `r length(unique(erd.iso$sp))` species for which ERD was determined (Fig.  S\@ref(fig:obs-comm-plc-curves).

@@

```{r erd-hyd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-hyd}Pairwise correlations between Effective Rooting Depth (ERD; mean $\\pm$ 1SE in m) and hydraulic properties for seven large, canopy species found on Barro Colorado Island (Panama). Scatterplots with linear fits are shown in the lower half. Corresponding Spearmon's $\\rho$ are given in the diagonally opposite location in the upper half, along with significance levels, where **, * and . represent *P*-value $\\le$ 0.01, 0.05 and 0.1, respectively. Diagonal shows density distribution of each variable.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/erd.stem.traits_cor.chart.jpeg"))
```

@@

### Future Directions for Model Development
\
The trade-offs that we identify between above- and below-ground traits, the water environment vertically below-ground and mortality rates are important for representing diversity in Dynamic Global Vegetation Models especially in species-rich forests. The hydraulic efficiency-safety trade-off between high values of maximum leaf or stem conductivity (high $K_{max}$) and cavitation-resistance (low $\Psi_{50}$ or $\Psi_{88}$) is hypothesized although community-wide examples are rare [@Hacke:2006; @Gleason.2016]. Such a trade-off is found at our study site at both leaf and stem level (Zailaa et al. to be submitted; Wolfe et al. in revision). We also found that LMAand WSG were strong predictors of PLC curves for leaves, but not for stems (not shown). Acquisitive traits such as high $K_{max}$ are expected to be associated with faster growth rates. We found a positive but non- significant relationship of adult growth rates ($\ge$ 10cm dbh) with $K_{max, ~leaf}$ (*R*^2^ = 0.13; *p* = 0.126, *n* = `r length(sp.exp.param$sp)`) and no relationship with $K_{max, ~stem}$ (*n* = `r stem.traits.sp` species). The relationships between ERD and above-ground hydraulic traits that we find will provide an important insight on how to model rooting depths and their coordination or trade-offs with other traits in order to better represent the functional diversity of tropical forests.

@@

```{r obs-comm-plc-curves, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "50%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-obs-comm-plc-curves}'Vulnerability curves' for species with ERD estimates. Curves in  data on ", length(sp.exp.param$sp)," tree species from BCI (A & B) with equation \\@ref(eq:vc-exp)--in absolute terms (A, B), as well as, in terms of loss of conductivity (%) (C, D). Each line represents a species, color coded by source of the curve: fitted to observed data (yellow ochre) on $K_{leaf}$ vs. $\\Psi_{leaf}$ or based on predicted model parameters $A$ and $B$ from soft traits. (gray).")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/kmax_by_psi/Leaf/obs.mod_plccurves_fits_ggplot_no_col.jpeg"))
```
